MCMC Informed Variational Bayes

### 2017-05-15 MIVB/MIVB.tex
code cleaning

\begin{figure}[h]
\centering
%\includegraphics{DLMAdvi}
\caption{The MCMC and ADVI posterior means for each $\boldsymbol{\theta}$ parameter with $100$ data replications at each $T \in \{50, 100, 250, 500, 1000, 2000, 5000, 10000\}$. True values for each parameter is denoted by the red line. MCMC consistently estimates of the true value, while both ADVI algorithms are only consistent for the mean parameter $\gamma$.
}
\label{fig:DLM:ADVI}
\end{figure}

A further $100 T = 500$ datasets are simulated with the previous parameter and hyperparameter values, and for each $\theta_i$ the ELBO is maximised for $q(\theta_i, | \boldsymbol{\lambda})$, holding each $\theta_{j \neq i}$ and $x_{0:T}$ fixed at their true values. $\boldsymbol{\lambda}$ includes only the mean and variance of $\theta_i$, so the ELBO is evaluated over a grid so the stochastic nature of the gradient ascent algorithm is not a confounding factor. The maximising value of the mean of $\theta_i$ is plotted in Figure \ref{fig:DLM:Numeric}. Holding $\sigma^2_x$ and $\sigma^2_y$ fixed at their true values improves the estimation of the mean of $\phi$, but neither $\sigma^2_y$ nor $\sigma^2_x$ improves.

\begin{figure}[h]
\centering
%\includegraphics{DLMNumeric}
\caption{The ELBO maximising mean for each $\theta$ parameter, holding each other $\theta$ and $x_{0:T}$ fixed. Estimation of the mean of $\phi$ improves relative to Figure \ref{fig:DLM:ADVI} but the posterior mean of $\sigma^2_y$ is consistently over-estimated while the posterior mean of $\sigma^2_x$ is consistently under-estimated.}
\label{fig:DLM:Numeric}
\end{figure}

@@

Treating the MCMC samples as `data', choosing a functional form for $q(\boldsymbol{\theta}, x_{1:T} | \boldsymbol{\lambda})$ can be approached as a traditional distribution fitting problem. Select optimal marginal distributions $\tilde{q}$ from a pool of suitable candidate distributions for each parameter by AIC, and for convenience select the same marginal distribution for each $x_t$ latent variable. A vine copula modelling approach is adopted for the dependence structure, as this can be selected independently to the marginal distribution. Sklar (1959) proves the any joint probability distribution with $k$ variables can be written as the product of the marginals and a copula function,
\begin{equation}
\label{vc1}
p(\omega_1, \dots, \omega_k) = p(\omega_1) \dots p(\omega_k) c(P(\omega_1), \dots, P(\omega_k))
\end{equation}
where $p(\theta)$ is a pdf, $P(\theta)$ is a cdf, and $c(\cdot)$ is a copula. A vine copula is a structure to factorise this copula, increasing the flexibility of modelling dependence, and is defined by a set of trees $\mathcal{V} = \{T_1, \dots, T_{k-1} \}$, where
\begin{enumerate}
\item $T_1$ has $k$ nodes corresponding to and labelled by the $k$ variables, with $k-1$ edges.
\item $T_t, t = 2, \dots, k-1$ have $k+1-t$ nodes corresponding to the edges in tree $T_{t-1}$
\item Two nodes in tree $T_t$ are joined by an edge only if the associated edges in $T_{t-1}$ share a node. 
\item Each node in $T_t,  t = 2, \dots, k-1$ is labelled by the union of the nodes associated edge in $T_{t-1}$ connects to.
\end{enumerate}
Each edge corresponds to a bivariate copula between the two variables that are unique to each of the connected nodes, with a conditioning set containing all shared variables. The product of each of these $k(k-1)/2$ bivariate copulae forms the complete copula. Figure (\ref{fig:vinecop}) provides an example for a four dimensional $\theta$.
\begin{figure}[h]
\centering
%\includegraphics[width=0.4\linewidth,height=\textheight,keepaspectratio]{vines.png}
\vspace{2mm}
\caption{An example of a Vine Copula over four variables. The edges in the top tree represent unconditional bivariate copulae between the parameters on the connected nodes. The second tree edges represent the bivariate copulae for $\theta_1 \theta_3 | \theta_2$ and for $\theta_3 \theta_4 | \theta_2$, the conditioning on $\theta_2$ caused by its appearance in each connected node. The edge in the final tree represents the bivariate copula for $\theta_1, \theta_4 | \theta_2, \theta_3$, as both $\theta_2$ and $\theta_3$ appear in the connected nodes.}
\label{fig:vinecop}
\end{figure}

Denoting the number of elements in $\boldsymbol{\theta}$ by $k$, a vine copula for $q(\boldsymbol{\theta}, x_{0:T} | \boldsymbol{\lambda})$ contains $T+k$ trees, with a number of parameters included in $\boldsymbol{\lambda}$ that is $O(T+k)^2$. Constructing the vine so that the first $k$ trees do not contain copulae between any two elements of $x_t$ and the $k+1$'th tree contains copulae $c(x_t, x_{t+1} | \boldsymbol{\theta})$ will force trees $k+r$, with $r \geq 2$, to contain copulae $c(x_t, x_{t+r} | \boldsymbol{\theta}, x_{t+1:t+r-1}). These can be set to the independence copula as the Markovian nature of the latent varables implies that $x_t$ and $x_{t+r}$ are conditionally independent given $\boldsymbol{\theta}$ and $x_{t+1:t+r-1}, reducing the number of copula parameters to $O(Tk)$. For convenience the vine is further restricted so that, in each of the first $k$ trees, every $x_t$ has an edge to the same $\theta$ node with the same copula family.

Di{\ss}mann's algorith (Di{\ss}mann, Brechmann, Czado and Kurowicka 2013) can be used to fit a vine copula to data, selecting both the set of edges for each tree and an appropriate copula family for each edge according to some criteria. A common critereon is to select the set of edges that maximise the sum of absolute empirical pairwise Kendell's Tau, with copula families that minimise AIC. This algorithm is modified to include the above restrictions. In the following, use of the letters $i$ and $j$ denote $\theta_i$ and $\theta_j$ and an alternative use of $t$ denotes $x_t$. Edges are denoted by $\{i, j | D \}$ and copula are given by $c(i, j | D)$ where $D$ is the conditioning set. Any copula that is to be included in the $q$ distribution is given by $\tilde{c}(i, j | D)$. Finally the empirical Kendell's Tau between a pair of variables is denoted by $\tau_{i, j | D}$. Noting that copulae are a function of the CDF of a random variable first transform all MCMC samples of $\boldsymbol{\theta}$ and $x_{0:T}$ to $(0, 1)$ using the empirical marginal CDFs. No restrictions are applied to the $\boldsymbol{\theta}$ sub-trees, so a $k$ dimensional vine for these parameters can be estimated using Di{\ss}mann's algorith by any critiera with selected edges $\{i, j | D\}$ and copulae $\tilde{c}(i, j | D)$ added to the first $k$ trees of the full vine.

The first tree also contains edges from each $x_t$ to the same $\theta$ element, selected by calculating $\tau_{j, t}$ for each $\theta_j$ and $x_t$ and add edges $\{t, i^{(1)} \}$ where
\begin{equation}
\label{Vine:TauCrit}
i^{(1)} = \arg \underset{j}{\max} \sum_{t} \tau_{j, t}.
\end{equation}
For each copula family $m$ and each $t$, estimate the copula $c_m(i^{(1)}, t)$ and choose an optimal family to include in the vine as $\tilde{c}(i^{(1)}, t)$. The optimal family could be chosen by criteria such as minimises the sum of AIC across $t$, the minimises AIC for the most $t$. In each tree $l = 2, \dots, k$ this processs is repeated: Calculate $\tau_{j, t | i^{(1)}, \dots, i^{(l-1)}$ for each $t$ and element of $\theta$ that has an edge to $\theta^{(l-1)}$ in one of the first $l-1$ trees, then choose $\theta_{i^{(l)}$ where
\begin{equation}
\label{Vine:TauCrit}
i^{(l)} = \arg \underset{j}{\max} \sum_{t} \tau_{j, t}.
\end{equation}
Estimate copulae $c_m(i^{(l)}, t | i^{(1:l-1)})$ for each $t$ and $m$ and select an optimal family as $\tilde{c}(i^{(l)}, t |i^{(1)}, \dots, i^{(l-1)})$.
Tree $k+1$ contains edges $\{t, t+1 |  i^{(1)}, \dots, i^{(k) \}$, with an optimal copula family $\tilde{c}(t, t+1 |i^{(1)}, \dots, i^{(l-k)})$. Tree $k+r$ with $r \geq 2$ must then contain edges $\{t, t+r |  i^{(1), \dots, i^{(k), t+1, \dots, t+r-1 \} with $\tilde{c}(t, t+r |  i^{(1), \dots, i^{(k), t+1, \dots, t+r-1) chosen to be the independence copula. This approach is given in Algorithm \ref{alg:MIVB}.

The product of all $\tilde{c}$ copula and selected $\tilde{q}$ marginals forms the approximating distribution $q(\boldsymbol{\theta}, x_{0:T} | \boldsymbol{\lambda})$. As the marginal distributions and copula for $x_t$ are identical for each $t$ additional latent variables $x_{T+1:T+S}$ can easily be included to form $q(\boldsymbol{\theta}, x_{0:T+S} | \boldsymbol{\lambda})$. The ELBO can be maximised for this $q$ using Copula Variational Inference (Tran, Blei and Airoldi 2015). As we are interested in forecasts of $y_{T+S+h}$ and $x_t$ is Markovian, the distribution of each $x_s, s < t$ is irrelevant given the distribution of $x_t$. Holding $\lambda}$ parameters assocaited with $x_t, t = 0, \dots, T$ fixed at MCMC data estimated values will drastically reduce the computational cost of the gradient ascent algorirhtm as only the $\lambda$ parameters associated with the marginal distributions of $\theta$ and $x_{T+1:T+S}$, as well as the copulae between these, needs to be optimised.

\begin{algorithm}[H]
 \SetKwInOut{Input}{Input}
 \Input{Observations of $\boldsymbol{\theta}, x_{1:T}$,  $k$, the number of elements in $\boldsymbol{\theta}$. \;}
 \KwResult{Set of copulae $\tilde{c}(\cdot, \codt)$}
 Transform observations for each variable to $(0, 1)$ by the inverse CDF\;
 Apply Di{\ss}mann's algorithm to $\boldsymbol{\theta}$, denote the selected edges by $\{i, j | D \}_l$ and copulae by $\tilde{c}(i, j | D)_l$ where $D$ is the conditioning set for each copula and $l = 1, \dots, k-1$ is the tree. \;
 Calculate $\hat{\tau}_{j, t}$ for each $j$ and $t$ \;
 Find $i^{(1)} = \arg \underset{j}{\max} \sum_t \hat{\tau}_{j, t}$. \;
 Add edges $\{i, j |\}_1$ and $\{i^{(1)}, t \}$ to Tree $1$. \;
 For each $t$ and copula family $m$ estimate $c_m(i^{(1)}, t)$.\;
 Select $\tilde{c}(i^{(1)}, t)$.\;
 For each $\tilde{c}(i, j)_1$ calculate $\theta_i | \theta_j$ and $\theta_j | \theta_i$; transform to (0, 1).\;
 For each $t$ use $\tilde{c}(i^{(1)}, t)$ to calculate $x_t | \theta_{i^{(1)}}$; transform to (0, 1).\;
 \For {$l = 2, \dots, k$} {
    Calculate $\hat{\tau}_{j, t | i^{(1:l-1)}}$ for $x_t | \theta_{i^{(1:l-1)}}$ and each $\theta_j | D$ where $D$ matches the $x_t$ conditioning set $\theta_{i^{(1:l-1)}}$. \;
  Find $i^{(l)} = \arg \underset{j}{\max} \sum_t \hat{\tau}_{j, t | i^{(1:l-1)}}$. \;
  Add edges $\{i, j | D\}_l$ and $\{i^{(l)}, t |  i^{(1:l-1)}\}$ to Tree $l$. \;
  For each $t$ and copula family $m$ estimate $c_m(i^{(l)}, t |i^{(1:l-1)})$.\;
  Select $\tilde{c}(i^{(l)}, t |i^{(1:l-1)})$.\;
  For each $\tilde{c}(i, j | D)_l$ calculate $\theta_i | \theta_j, \theta_D$ and $\theta_j | \theta_i, \theta_D$; transform to (0, 1).\;
  For each $\tilde{c}(i^{(l)}, t |i^{(1:l-1)})$ calculate $x_t | \theta_{i^{(1:l)}}$; transform to (0, 1).\;
 }
  For each $t$ in $2, \dots, T$ and copula family $m$ estimate $c_m(t, t-1 | \boldsymbol{\theta}).$ \;
  Select $\tilde{c}(t, t-1 | \boldsymbol{\theta})$. \;
   \caption{MIVB Vine selection algorithm}
  \label{alg:MIVB}
\end{algorithm}

>>>>

\begin{figure}[h]
\includegraphics[width=1\linewidth]{DLMAdvi}
\caption{The MCMC and ADVI posterior means for each $\boldsymbol{\theta}$ parameter with $100$ data replications at each $T \in \{50, 100, 250, 500, 1000, 2000\}$. True values for each parameter is denoted by the red line. MCMC consistently estimates of the true value, while both ADVI algorithms are only consistent for the mean parameter $\gamma$.}
\label{fig:DLM:ADVI}
\end{figure}

A further $100 T = 500$ datasets are simulated with the previous parameter and hyperparameter values, and for each $\theta_i$ the ELBO is maximised for $q(\theta_i, | \boldsymbol{\lambda})$, holding each $\theta_{j \neq i}$ and $x_{0:T}$ fixed at their true values, and variance of $\theta_i$ set at the MCMC sample variance. $\lambda$ contains only the mean of $\theta_i$, so the ELBO is evaluated over a grid so the stochastic nature of the gradient ascent algorithm is not a confounding factor. The maximising value of the mean of $\theta_i$ is plotted in Figure \ref{fig:DLM:Numeric}. Holding $\sigma^2_x$ and $\sigma^2_y$ fixed at their true values improves the estimation of the mean of $\phi$, but neither $\sigma^2_y$ nor $\sigma^2_x$ improves.

\begin{figure}[h]
\centering
\includegraphics{DLMNumeric}
\caption{The ELBO maximising mean for each $\theta$ parameter, holding each other $\theta$ and $x_{0:T}$ fixed at their true values. Estimation of the mean of $\phi$ and $\sigma^2_x$ improves relative to Figure \ref{fig:DLM:ADVI} but the posterior mean of $\sigma^2_y$ is consistently over-estimated.}
\label{fig:DLM:Numeric}
\end{figure}

@@

Treating the MCMC samples as `data', choosing a functional form for $q(\boldsymbol{\theta}, x_{1:T} | \boldsymbol{\lambda})$ can be approached as a traditional distribution fitting problem. Select optimal marginal distributions $\tilde{q}$ from a pool of suitable candidate distributions for each parameter by AIC, and for convenience select the same marginal distribution for each $x_t$ latent variable. A vine copula modelling approach is adopted for the dependence structure, as this can be selected independently to the marginal distribution. Sklar (1959) proves the any joint probability distribution with $k$ variables can be written as the product of the marginals and a copula function,
\begin{equation}
\label{vc1}
p(\omega_1, \dots, \omega_k) = p(\omega_1) \dots p(\omega_k) c(P(\omega_1), \dots, P(\omega_k))
\end{equation}
where $p(\theta)$ is a pdf, $P(\theta)$ is a cdf, and $c(\cdot)$ is a copula. A vine copula is a structure to factorise this copula, increasing the flexibility of modelling dependence, and is defined by a set of trees $\mathcal{V} = \{T_1, \dots, T_{k-1} \}$, where
\begin{enumerate}
\item $T_1$ has $k$ nodes corresponding to and labelled by the $k$ variables, with $k-1$ edges.
\item $T_t, t = 2, \dots, k-1$ have $k+1-t$ nodes corresponding to the edges in tree $T_{t-1}$
\item Two nodes in tree $T_t$ are joined by an edge only if the associated edges in $T_{t-1}$ share a node. 
\item Each node in $T_t,  t = 2, \dots, k-1$ is labelled by the union of the nodes associated edge in $T_{t-1}$ connects to.
\end{enumerate}
Each edge corresponds to a bivariate copula between the two variables that are unique to each of the connected nodes, with a conditioning set containing all shared variables. The product of each of these $k(k-1)/2$ bivariate copulae forms the complete copula. Figure (\ref{fig:vinecop}) provides an example for a four dimensional $\theta$.
\begin{figure}[h]
\centering
\includegraphics[width=0.4\linewidth,height=\textheight,keepaspectratio]{vines.png}
\vspace{2mm}
\caption{An example of a Vine Copula over four variables. The edges in the top tree represent unconditional bivariate copulae between the parameters on the connected nodes. The second tree edges represent the bivariate copulae for $\theta_1 \theta_3 | \theta_2$ and for $\theta_3 \theta_4 | \theta_2$, the conditioning on $\theta_2$ caused by its appearance in each connected node. The edge in the final tree represents the bivariate copula for $\theta_1, \theta_4 | \theta_2, \theta_3$, as both $\theta_2$ and $\theta_3$ appear in the connected nodes.}
\label{fig:vinecop}
\end{figure}

Denoting the number of elements in $\boldsymbol{\theta}$ by $k$, a vine copula for $q(\boldsymbol{\theta}, x_{0:T} | \boldsymbol{\lambda})$ contains $T+k$ trees, with a number of parameters included in $\boldsymbol{\lambda}$ that is $O(T+k)^2$. Constructing the vine so that the first $k$ trees do not contain copulae between any two elements of $x_t$ and the $k+1$'th tree contains copulae $c(x_t, x_{t+1} | \boldsymbol{\theta})$ will force trees $k+r$, with $r \geq 2$, to contain copulae $c(x_t, x_{t+r} | \boldsymbol{\theta}, x_{t+1:t+r-1})$. These can be set to the independence copula as the Markovian nature of the latent varables implies that $x_t$ and $x_{t+r}$ are conditionally independent given $\boldsymbol{\theta}$ and $x_{t+1:t+r-1}$, reducing the number of copula parameters to $O(Tk)$. For convenience the vine is further restricted so that, in each of the first $k$ trees, every $x_t$ has an edge to the same $\theta$ node with the same copula family.

Di{\ss}mann's algorith (Di{\ss}mann, Brechmann, Czado and Kurowicka 2013) can be used to fit a vine copula to data, selecting both the set of edges for each tree and an appropriate copula family for each edge according to some criteria. A common critereon is to select the set of edges that maximise the sum of absolute empirical pairwise Kendall's Tau, with copula families that minimise AIC. This algorithm is modified to include the above restrictions. In the following, use of the letters $i$ and $j$ denote $\theta_i$ and $\theta_j$ and an alternative use of $t$ denotes $x_t$. Edges are denoted by $\{i, j | D \}$ and copula are given by $c(i, j | D)$ where $D$ is the conditioning set. Any copula that is to be included in the $q$ distribution is given by $\tilde{c}(i, j | D)$. Finally the empirical Kendall's Tau between a pair of variables is denoted by $\tau_{i, j | D}$. Noting that copulae are a function of the CDF of a random variable first transform all MCMC samples of $\boldsymbol{\theta}$ and $x_{0:T}$ to $(0, 1)$ using the empirical marginal CDFs. No restrictions are applied to the $\boldsymbol{\theta}$ sub-trees, so a $k$ dimensional vine for these parameters can be estimated using Di{\ss}mann's algorith by any critiera with selected edges $\{i, j | D\}$ and copulae $\tilde{c}(i, j | D)$ added to the first $k$ trees of the full vine.

The first tree also contains edges from each $x_t$ to the same $\theta$ element, selected by calculating $\tau_{j, t}$ for each $\theta_j$ and $x_t$ and add edges $\{t, i^{(1)} \}$ where
\begin{equation}
\label{Vine:TauCrit}
i^{(1)} = \arg \underset{j}{\max} \sum_{t} \tau_{j, t}.
\end{equation}
For each copula family $m$ and each $t$, estimate the copula $c_m(i^{(1)}, t)$ and choose one family for all of these copula as $\tilde{c}(i^{(1)}, t)$. The decision could be on the basis of criteria such as `minimises the sum of AIC across $t$' or `minimises AIC for the most $t$'. In each tree $l = 2, \dots, k$ this processs is repeated: Calculate $\tau_{j, t | i^{(1)}, \dots, i^{(l-1)}}$ for each $t$ and element of $\theta$ that has an edge to $\theta^{(l-1)}$ in one of the first $l-1$ trees, then choose $\theta_{i^{(l)}}$ where
\begin{equation}
\label{Vine:TauCrit}
i^{(l)} = \arg \underset{j}{\max} \sum_{t} \tau_{j, t}.
\end{equation}
Estimate copulae $c_m(i^{(l)}, t | i^{(1:l-1)})$ for each $t$ and $m$ and select a family as $\tilde{c}(i^{(l)}, t |i^{(1:l-1)})$. Tree $k+1$ contains edges $\{t, t+1 |  i^{(1:k)} \}$, with an optimal copula family $\tilde{c}(t, t+1 |i^{(1:k)})$. Tree $k+r$ with $r > 1$ must then contain edges $\{t, t+r |  i^{(1:k)}, t+1, \dots, t+r-1 \}$ with $\tilde{c}(t, t+r |  i^{(1:k)}, t+1, \dots, t+r-1)$ chosen to be the independence copula. This approach is given in Algorithm \ref{alg:MIVB}.

The product of all $\tilde{c}$ copula and selected $\tilde{q}$ marginals forms the approximating distribution $q(\boldsymbol{\theta}, x_{0:T} | \boldsymbol{\lambda})$. As the marginal distributions and copula for $x_t$ are identical for each $t$ additional latent variables $x_{T+1:T+S}$ can easily be included to form $q(\boldsymbol{\theta}, x_{0:T+S} | \boldsymbol{\lambda})$. The ELBO can be maximised for this $q$ using Copula Variational Inference (Tran, Blei and Airoldi 2015). As we are interested in forecasts of $y_{T+S+h}$ and $x_t$ is Markovian, the distribution of each $x_t$ is irrelevant given the distribution of $x_s, s > t$. Holding $\lambda$ parameters assocaited with $x_t, t = 0, \dots, T$ fixed at MCMC data estimated values will drastically reduce the computational cost of the gradient ascent algorirhtm as only the $\lambda$ parameters associated with the marginal distributions of $\theta$ and $x_{T+1:T+S}$, as well as the copulae between these, needs to be optimised.

\begin{algorithm}[H]
 \SetKwInOut{Input}{Input}
 \Input{Observations of $\boldsymbol{\theta}, x_{1:T}$,  $k$, the number of elements in $\boldsymbol{\theta}$.}
 \KwResult{Set of copulae $\tilde{c}(\cdot, \cdot)$}
 Transform observations for each variable to $(0, 1)$ by the inverse CDF\;
 Apply Di{\ss}mann's algorithm to $\boldsymbol{\theta}$, denote the selected edges by $\{i, j | D \}_l$ and copulae by $\tilde{c}(i, j | D)_l$ where $D$ is the conditioning set for each copula and $l = 1, \dots, k-1$ is the tree. \;
 Calculate $\hat{\tau}_{j, t}$ for each $j$ and $t$ \;
 Find $i^{(1)} = \arg \underset{j}{\max} \sum_t \hat{\tau}_{j, t}$. \;
 Add edges $\{i, j |\}_1$ and $\{i^{(1)}, t \}$ to Tree $1$. \;
 For each $t$ and copula family $m$ estimate $c_m(i^{(1)}, t)$.\;
 Select $\tilde{c}(i^{(1)}, t)$.\;
 For each $\tilde{c}(i, j)_1$ calculate $\theta_i | \theta_j$ and $\theta_j | \theta_i$; transform to (0, 1).\;
 For each $t$ use $\tilde{c}(i^{(1)}, t)$ to calculate $x_t | \theta_{i^{(1)}}$; transform to (0, 1).\;
 \For {$l = 2, \dots, k$} {
    Calculate $\hat{\tau}_{j, t | i^{(1:l-1)}}$ for $x_t | \theta_{i^{(1:l-1)}}$ and each $\theta_j | D$ where $D$ matches the $x_t$ conditioning set $\theta_{i^{(1:l-1)}}$. \;
  Find $i^{(l)} = \arg \underset{j}{\max} \sum_t \hat{\tau}_{j, t | i^{(1:l-1)}}$. \;
  Add edges $\{i, j | D\}_l$ and $\{i^{(l)}, t |  i^{(1:l-1)}\}$ to Tree $l$. \;
  For each $t$ and copula family $m$ estimate $c_m(i^{(l)}, t |i^{(1:l-1)})$.\;
  Select $\tilde{c}(i^{(l)}, t |i^{(1:l-1)})$.\;
  For each $\tilde{c}(i, j | D)_l$ calculate $\theta_i | \theta_j, \theta_D$ and $\theta_j | \theta_i, \theta_D$; transform to (0, 1).\;
  For each $\tilde{c}(i^{(l)}, t |i^{(1:l-1)})$ calculate $x_t | \theta_{i^{(1:l)}}$; transform to (0, 1).\;
 }
  For each $t$ in $2, \dots, T$ and copula family $m$ estimate $c_m(t, t-1 | \boldsymbol{\theta}).$ \;
  Select $\tilde{c}(t, t-1 | \boldsymbol{\theta})$. \;
   \caption{MIVB Vine selection algorithm}
  \label{alg:MIVB}
\end{algorithm}
